%%%%%%%%%%%%%%%%
\section{Thursday 23 August:  Raster Display v/s Vector Display}
%%%%%%%%%%%%%%%%%

\

{\bf Ray Tracing}:  Which objects intersect the line?

\

{\bf Polygon Projection (Polygon Rendering)}

Triangles on the screen

Vertex coordinates

Multiply by matrices to convert to screen's coordinate system

\

Modeling Coordinates 

\qquad $\to$ World Coordinates

\qquad $\to$ Viewing and Projection Coordinates (Camera View)

\qquad $\to$ Normalized Coordinates

\qquad $\to$ Device Coordinates

\

\begin{tabular}{m{0.3in}m{.2in}*4{m{1.0in}m{.2in}}}
	$\left[ \begin{matrix} x \cr y \cr z \cr w \cr \end{matrix}\right]$
	&
	$\to$
	&
	Model \par Martrix
	&
	$\to$
	&
	Projection \par Matrix \par (Camera View)
	&
	$\to$
	&
	Perspective \par Division
	&
	$\to$
	&
	Viewpoint \par Transformation
	\cr
\end{tabular}

\

{\bf 2D Scan Conversion}:  Converts 2D object description to pixel values. 

{\it ``Pixel''}  Picture Element

Visible Surface Determination:  Occlusion

\

Direct Illumination v/s Global Illumination:  Shadows not completely dark

Curves and surfaces give a smooth alternative to polygonal representation of objects.  

\

Lots of student projects have dealt with {\bf Particle Dynamics}

\

{\bf Early Video Games}  Nimatron (1940)  First ``video'' game used an oscilloscope as its screen.  

\

{\bf Raster Display} - Image broken into pixels

{\bf Vector Display} - Smooth curves, like moving lasers or electron beam.  

\

Black \& white {\bf bitmap} has one bit per pixel

\

{\bf Raster Displays}

More computationally expensive

Requires more memory

Constant refresh rate

Supports area fills

Won over vector after memory got cheap.  

\

Frame buffers are getting more complex.  

\qquad Double buffering so screen doesn't refresh in the middle of a memory move

\qquad Left and right buffers for stereoscopic rendering

\qquad Depth buffer for occlusion

\

{\bf Know from Today's Lesson}

Pixel, raster, bitmap, frame buffer, {\it aliasing}

Know the difference between raster and vector.  

\

{\bf Aliasing} - in computer graphics, the jagged, or saw-toothed appearance of curved or diagonal lines on a low-resolution monitor.

